The Categorical Structure of Profunctor Optics

### 2017-11-06 paper.tex
Get junk out of proof

@@

\begin{proposition}
\label{prop-profunctor-optics-are-optics}
\begin{align*}
[\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) &\cong \PreOptic_\M((S, S), (A, A))
\end{align*}
\end{proposition}
\begin{proof}
We have the chain of isomorphisms:
\begin{align*}
&[\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) \\
\cong \;&[\Tamb_\C(\M), \Set](\Tamb_\C(\M)(\Pastro_\M E_A, -), (U-)(S,S))  && \text{(by Lemma \ref{lemma-rep})}\\
\cong \;&(U\Pastro_\M E_A)(S,S)  && \text{(by Yoneda)} \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S) \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times \C(C,A) \times \C(A,D) \times \C(MD,S) \\
\cong \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S)  && \text{(by co-Yoneda twice)} \\
= \;&\PreOptic_\M((S, S), (A, A))
\end{align*}
\todo{This works for $\PreOptic_\M((S, S'), (A, A'))$ just fine, just define $E_{A,A'}$ in the obvious way.}

@@

\begin{proposition}
If $p : S \hto A$ is an optic then the associated natural transformation $(U-)(A,A) \to (U-)(S,S)$ is a lax monoidal.
\end{proposition}
\begin{proof}
The final few isomorphisms used in Proposition \ref{prop-profunctor-optics-are-optics} are essentially window dressing. Suppose we have an element $\langle l, e_1, e_2, r \rangle$ of 
\begin{align*}
(U\Pastro_\M E_A)(S,S) = \int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S).
\end{align*}
The round trip through the isomorphism $\Phi : (U\Pastro_\M E_A)(S,S) \cong \PreOptic_\M((S, S), (A, A))$ demonstrates that $\langle l, e_1, e_2, r \rangle = \langle (M e_1)l, \id_A, \id_A, r(M e_2)\rangle$. Hence we may always assume we have a representative of the form $\langle l, \id_A, \id_A, r \rangle$. This representative corresponds to an optic via the isomorphism $\Phi$ if both $rl = \id_S$ and $\langle lr\rangle = \langle \phi A \rangle$ in $\int^{M \in \M} \C(M A, M A)$. Note that just as we can ``strictify'' a representative of an optic, if we have an element of $(U\Pastro_\M E_A)(S,S)$ that corresponds to an optic we may always choose a representative so that the latter identity holds on the nose.

Next we examine the image of the optics through the Yoneda isomorphism \[ (U\Pastro_\M E_A)(S,S) \cong [\Tamb_\C(\M), \Set](\Tamb_\C(\M)(\Pastro_\M E_A, -), (U-)(S,S)) \] 
By the definition of the Yoneda isomorphism, a natural transformation \[ p : \Tamb_\C(\M)(\Pastro_\M E_A, -) \to (U-)(S,S) \] will correspond to an optic iff the image of the identity morphism $\id : \Pastro_\M E_A \to \Pastro_\M E_A$ through the component $p_{\Pastro_\M E_A}$ is an optic in $(U\Pastro_\M E_A)(S, S)$.

Finally the isomorphism \[ [\Tamb_\C(\M), \Set](\Tamb_\C(\M)(\Pastro_\M E_A, -), (U-)(S,S)) \cong \; [\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) \] is given by precomposition by the isomorphism of Lemma \ref{lemma-rep}, so in summary, a ``profunctor preoptic'' $p : (U-)(A, A) \to (U-)(S,S)$ is a lawful optic iff $p_{\Pastro_\M E_A}(\langle \id_A, \id_A, \id_A, \id_A \rangle)$ is an optic in $(U\Pastro_\M E_A)(S, S)$.

To show $p$ is monoidal, we must show that both diagrams
\[
\begin{tikzcd}
UP(A,A) \times UQ(A,A) \ar[r, "p_P \times p_Q"] \ar[d, swap] & UP(S, S) \times UQ(S, S)  \ar[d] \\
U(P \odot Q)(A,A)  \ar[r, "p_{P \odot Q}", swap] & U(P \odot Q)(S,S)
\end{tikzcd}
\]
and
\[
\begin{tikzcd}
1 \ar[d] \ar[dr] & \\
UI(A,A)  \ar[r, "p_I", swap] & UI(S,S)
\end{tikzcd}
\]
commute. For the first we have
\begin{align*}
\langle (p_P \times p_Q)(x, y) \rangle &= \langle (UP)(l,r) \alpha_{A,A,M} x,  (UQ)(l,r) \beta_{A,A,M} y\rangle \\
&= \langle (UP)(l,lr) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations)} \\
&= \langle (UP)(l,\phi A) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$)} \\
&= \langle (UP)((\phi A)l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations again)} \\
&= \langle (UP)(lrl,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$ again)} \\
&= \langle (UP)(l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $rl = \id_S$)} \\
&= (U(P \odot Q))(l,r) \langle \alpha_{A,A,M} x, \beta_{A,A,M} y \rangle && \text{(by definition of $P \odot Q$)} \\
&= (U(P \odot Q))(l,r) (\alpha_{A,A,M} \odot \beta_{A,A,M}) \langle x, y \rangle && \text{(by definition of $\alpha \odot \beta$)} \\
&= p_{P \odot Q} \langle x, y \rangle
\end{align*}
And for the second:
\begin{align*}
p_I(\id_A) = (UI)(l,r) (M \id_A) = r \id_{MA} l = rl = \id_S
\end{align*}
as required.
\end{proof}

>>>>

\begin{proposition}
The category $\PreOptic_\M$ is isomorphic to the full subcategory on objects of the form $E_A$ in the Kleisli category for the monad $U \Pastro_\M : \Prof(\C, \C) \to \Prof(\C, \C)$.
\end{proposition}
\begin{proof}
\todo{Is there a point to proving this?}
\end{proof}

@@

\begin{proposition}
\label{prop-profunctor-optics-are-optics}
\begin{align*}
[\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) &\cong \PreOptic_\M(S, A)
\end{align*}
\end{proposition}
\begin{proof}
We have the chain of isomorphisms:
\begin{align*}
&[\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) \\
\cong \;&[\Tamb_\C(\M), \Set](\Tamb_\C(\M)(\Pastro_\M E_A, -), (U-)(S,S))  && \text{(by Lemma \ref{lemma-rep})}\\
\cong \;&(U\Pastro_\M E_A)(S,S)  && \text{(by Yoneda)} \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S) \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times \C(C,A) \times \C(A,D) \times \C(MD,S) \\
\cong \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S)  && \text{(by co-Yoneda twice)} \\
= \;&\PreOptic_\M(S, A)
\end{align*}
\todo{This works for $\PreOptic_\M((S, S'), (A, A'))$ just fine, just define $E_{A,A'}$ in the obvious way.}

@@

\begin{proposition}
If $p : S \hto A$ is an optic then the associated natural transformation $(U-)(A,A) \to (U-)(S,S)$ is a lax monoidal. \todo{Rename the natural transformation $p$}
\end{proposition}
\begin{proof}
Let $\langle l, r \rangle$ be a representative for $p$ in $\PreOptic_\M(S,A)$ via the isomorphism of Proposition \ref{prop-profunctor-optics-are-optics}, so that
\begin{align*}
p_P = (UP)(l,r) \alpha_{A,A,M}
\end{align*}
on each Tambara module $P$.

To show $p$ is monoidal, we must show that both
\[
\begin{tikzcd}
UP(A,A) \times UQ(A,A) \ar[r, "p_P \times p_Q"] \ar[d, swap] & UP(S, S) \times UQ(S, S)  \ar[d] \\
U(P \odot Q)(A,A)  \ar[r, "p_{P \odot Q}", swap] & U(P \odot Q)(S,S)
\end{tikzcd}
\]
and
\[
\begin{tikzcd}
1 \ar[d] \ar[dr] & \\
UI(A,A)  \ar[r, "p_I", swap] & UI(S,S)
\end{tikzcd}
\]
commute. For the first we have
\begin{align*}
\langle (p_P \times p_Q)(x, y) \rangle &= \langle (UP)(l,r) \alpha_{A,A,M} x,  (UQ)(l,r) \beta_{A,A,M} y\rangle \\
&= \langle (UP)(l,lr) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations)} \\
&= \langle (UP)(l,\phi A) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$)} \\
&= \langle (UP)((\phi A)l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations again)} \\
&= \langle (UP)(lrl,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$ again)} \\
&= \langle (UP)(l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $rl = \id_S$)} \\
&= (U(P \odot Q))(l,r) \langle \alpha_{A,A,M} x, \beta_{A,A,M} y \rangle && \text{(by definition of $P \odot Q$)} \\
&= (U(P \odot Q))(l,r) (\alpha_{A,A,M} \odot \beta_{A,A,M}) \langle x, y \rangle && \text{(by definition of $\alpha \odot \beta$)} \\
&= p_{P \odot Q} \langle x, y \rangle
\end{align*}
And for the second:
\begin{align*}
p_I(\id_A) = (UI)(l,r) (M \id_A) = r \id_{MA} l = rl = \id_S
\end{align*}
as required.
\end{proof}
